Spotify Data

### 2024-10-10 paper/paper.qmd
Update

---
title: "Toronto Beach Data"
subtitle: "Data Analysis of the Relationship Between Date of Observation, Air Temperatures, Water Temperatures, and Waterfowl Activities"
author: "Yanzun Jiang"
date: today
date-format: long
thanks: "Code and data supporting this analysis is available at: <https://github.com/Stary54264/Toronto-Beach-Data>"
abstract: "This report analyzes environmental data collected in the summer of 2023 regarding Toronto's beaches, focusing on variables such as air temperature, water temperature, and waterfowl observations. The findings reveal that air temperatures have a linear relationship with water temperatures, while increased water temperatures correlate with higher waterfowl activity in some content. These insights emphasize the importance of understanding how weather patterns affect wildlife behavior and beach usability, suggesting that public awareness and management strategies should be adapted accordingly. Future research should explore the impact of these environmental factors on beachgoer experiences and the broader ecological implications."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

In 2014, Toronto experienced an extremely cold summer that impacted its beaches (@torontoweather). During this period, data on key environmental factors, such as air temperature, water temperature, and observations of waterfowl activity, was collected at various beaches across the city (@opendatatoronto). These factors are critical as they directly influence the ecological balance of the beaches, the safety of recreational activities, and the overall enjoyment of beachgoers (@weatherbeach).

Consequently, analyzing this environmental data is essential to understanding how these variables affect the behavior of waterfowl and the suitability of the beaches for public use. According to @temperaturewaterfowl, waterfowl presence can be influenced by temperature changes, precipitation, and human activity on the beaches. This study examines the relationships between air temperature, water temperature, and waterfowl observations, and how these conditions might correlate with fluctuations in bird activity and potential impacts on beachgoers.

To conduct this analysis, a dataset collected by @opendatatoronto was utilized, as described in @sec-data. Based on the initial findings, it was observed that higher air temperatures are correlated with higher water temperatures, while higher water temperatures coincided with an increase in waterfowl activity in some extent(@sec-results). @sec-discussion explores the key findings from the analysis, highlighting significant seasonal patterns, correlations between variables, and identifying areas for future research and management strategies to enhance ecosystem health and public safety. Further recommendations include focusing on incorporating year-round data and addressing NA values to enhance the analysis and insights on Toronto beaches. This paper's structure includes an overview of the data and methods in @sec-data, the results of the analysis in @sec-results, a detailed discussion of the results in @sec-discussion, and supplementary insights in @sec-appendix.

@@

The dataset used in this analysis is the "Toronto Beaches Observations" sourced from Toronto's Open Data platform (@opendatatoronto). It records various aspects of the beaches in Toronto from 2008 to present. In Toronto, water quality at beaches is regularly monitored to ensure public safety, and environmental variables, such as air and water temperature, and the presence of waterfowl, are tracked for ecological and recreational purposes (@opendatatoronto).

The variables analyzed in this study include "Date," which records the day each measurement was taken; "Air Temperature," which represents the temperature of the atmosphere at the beach site; "Water Temperature," indicating the temperature of the water in designated swimming areas; and "Waterfowl Observations," which track the number of birds spotted near the water (@opendatatoronto).

The dataset was accessed using the `opendatatoronto` package (@opendatatoronto). For the analysis, the R programming language was employed (@citeR), utilizing the `tidyverse` (@citetidyverse) package for data cleaning, transformation, visualization, and the `janitor` package (@citejanitor) for initial cleaning and formatting of the raw data. Afterward, the cleaned dataset was processed and tested using additional functions from the `tidyverse` package (@citetidyverse).

## Measurement

Each variable in this dataset was measured with specific techniques that ensure accuracy and relevance to the analysis of Toronto beaches.

Air temperatures were measured using calibrated thermometers placed in shaded areas to avoid direct sunlight, providing accurate readings of the ambient temperature. Measurements were taken at regular intervals throughout the day, typically during peak observation hours, to capture variations in temperature across different times (@measurement).

Water temperature was recorded using submerged thermometers positioned at varying depths. This approach allowed for capturing temperature gradients that could occur in different water layers, offering insights into thermal stratification, especially during warmer months (@measurement).

Observations of waterfowl were conducted using a standardized counting method and estmating method. Trained observers recorded the number of birds present at various times throughout the day, ensuring that counts were conducted under similar conditions to minimize variability. Then they use the samples to estimate the total number of waterfowls. This method also included recording species, which adds valuable information about the biodiversity present in the area (@opendatatoronto).

By discussing these measurement techniques, we establish a clear connection between the dataset and the reliability of the results. This thorough exploration of measurement practices highlights the importance of consistent and accurate data collection in understanding the environmental dynamics of Toronto beaches.

@@

The preview of the dataset (@tbl-data) spans from 2008 to present. The dataset provides a comprehensive overview of key environmental variables observed at Toronto's beaches, including the date of observation, air and water temperatures, and waterfowl activity. These variables are crucial for understanding long-term trends in environmental conditions and their impact on beach safety, water quality, and wildlife behavior. This preview serves as an introduction to the dataset's structure and quality, setting the stage for in-depth analysis and trend exploration over time.

```{r}
#| label: tbl-data
#| tbl-cap: Preview of Data
#| echo: false
#| warning: false
#| message: false

kable(head(data), format = "markdown")
```

## Data Visualization

### Distribution of the Variables

These histograms (@fig-air, @fig-water, and @fig-fowl) visualizes the distribution of air temperatures, water temperature, and waterfowl observations across the dataset from 2008 to the present. These graphs helps identify patterns, outliers, and the overall spread of the data. By examining these distributions, we can gain insights into how they behave over time and how it may influence other variables, such as water quality or wildlife observations.

```{r}
#| label: fig-air
#| fig-cap: Distribution of Air Temperature at Toronto Beaches
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = air_temp)) +
  geom_histogram(bins = 30, col = "darkblue", fill = "lightblue") +
  theme_minimal() +
  labs(x = "Air Temperature (°C)", y = "Frequency")
```

```{r}
#| label: fig-water
#| fig-cap: Distribution of Water Temperature at Toronto Beaches
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = water_temp)) +
  geom_histogram(bins = 30, col = "darkblue", fill = "lightblue") +
  theme_minimal() +
  labs(x = "Water Temperature (°C)", y = "Frequency")
```

```{r}
#| label: fig-fowl
#| fig-cap: Distribution of Waterfowl Observations at Toronto Beaches
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = water_fowl)) +
  geom_histogram(bins = 30, col = "darkblue", fill = "lightblue") +
  theme_minimal() +
  labs(x = "Number of Waterfowl Observed", y = "Frequency")
```

### Changes of Beaches with Time

These scatter plots (@fig-year-air, @fig-year-water, and @fig-year-fowl) visualize the relationship between each environmental variable and time, spanning from 2008 to the present. By plotting air temperature, water temperature, and waterfowl observations against time, we aim to identify trends, and potential long-term changes in these variables at Toronto's beaches. Each scatter plot provides insights into how these environmental factors have fluctuated over the years, revealing trends in beach weather conditions and wildlife presence. These time-based visualizations will aid in understanding any periodic or gradual changes and their potential impact on beach safety and ecological health.

```{r}
#| label: fig-year-air
#| fig-cap: Air Temperature Over Years at Toronto Beaches
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = year, y = air_temp)) +
  geom_point(col = "lightblue") +
  theme_minimal() +
  labs(x = "Year", y = "Air Temperature (°C)")
```

```{r}
#| label: fig-year-water
#| fig-cap: Water Temperature Over Years at Toronto Beaches
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = year, y = water_temp)) +
  geom_point(col = "lightblue") +
  theme_minimal() +
  labs(x = "Year", y = "Water Temperature (°C)")
```

```{r}
#| label: fig-year-fowl
#| fig-cap: Waterfowl Observations Over Years at Toronto Beaches
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = year, y = water_fowl)) +
  geom_point(col = "lightblue") +
  theme_minimal() +
  labs(x = "Year", y = "Number of Waterfowl Observed")
```

These scatter plots (@fig-month-air, @fig-month-water, and @fig-month-fowl) illustrate the seasonal variations in air temperature, water temperature, rainfall occurrence, and waterfowl observations at Toronto beaches from 2008 to the present. By plotting these variables over time, distinct seasonal patterns emerge. The scatter plots help visualize how these environmental factors fluctuate throughout the year, reflecting natural seasonal cycles. Observing these recurring trends provides valuable insights into the seasonal dynamics affecting beach ecosystems and recreational conditions.

```{r}
#| label: fig-month-air
#| fig-cap: Seasonal Change of Air Temperature at Toronto Beaches
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = month, y = air_temp)) +
  geom_point(col = "lightblue") +
  theme_minimal() +
  labs(x = "Month", y = "Air Temperature (°C)")
```

```{r}
#| label: fig-month-water
#| fig-cap: Seasonal Change of Water Temperature at Toronto Beaches
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = month, y = water_temp)) +
  geom_point(col = "lightblue") +
  theme_minimal() +
  labs(x = "Month", y = "Water Temperature (°C)")
```

```{r}
#| label: fig-month-fowl
#| fig-cap: Seasonal Change of Waterfowl Observations at Toronto Beaches
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = month, y = water_fowl)) +
  geom_point(col = "lightblue") +
  theme_minimal() +
  labs(x = "Month", y = "Number of Waterfowl Observed")
```

### Correlations Between Observed Variables

These scatter plots (@fig-air-water, @fig-water-fowl) explore the relationships between various observed variables, such as air temperature, water temperature, and waterfowl observations, to identify potential correlations. By plotting some variables against some other, we aim to uncover how these environmental factor might influence another. For instance, the relationship between air temperature and water temperature may reveal patterns in heat transfer between the atmosphere and water, while the correlation between water temperature and waterfowl observations could indicate how weather conditions affect wildlife activity. These scatter plots provide a deeper understanding of the interconnectedness of these variables and offer insights into how environmental conditions at Toronto beaches influence each other.

```{r}
#| label: fig-air-water
#| fig-cap: Relationship Between Air Temperature and Water Temperature at Toronto Beaches
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = air_temp, y = water_temp)) +
  geom_point(col = "lightblue") +
  theme_minimal() +
  labs(x = "Air Temperature (°C)", y = "Water Temperature (°C)")
```

```{r}
#| label: fig-water-fowl
#| fig-cap: Relationship Between Water Temperature and Waterfowl Observations at Toronto Beaches
#| echo: false
#| warning: false
#| message: false

data |>
  ggplot(aes(x = water_temp, y = water_fowl)) +
  geom_point(col = "lightblue") +
  theme_minimal() +
  labs(x = "Water Temperature (°C)", y = "Number of Waterfowl Observed")
```

@@

## Gain from Changes of Beaches with Time

@fig-year-air shows the relationship between air temperature and time. Over the years, we observe a decrease followed by an increase in air temperatures, with potential gradual trends reflecting long-term changes rather than distinct seasonal cycles. This could indicate climate-related variations or other long-term environmental shifts affecting the air temperature at the beaches.

@fig-year-water displays water temperature versus time. Similar to air temperature, water temperature shows fluctuations across years, but with a less explicit pattern. The data suggests that the overall trends could be linked to climate change or weather anomalies.

@fig-year-fowl illustrates waterfowl observations over time. The scatter plot reveals variations in bird activity across different years. While no sharp seasonal patterns are apparent due to the yearly scale, the data might point toward longer-term changes in waterfowl populations, potentially related to environmental factors, habitat changes, or migratory patterns.

These scatter plots (@fig-year-air, @fig-year-water, and @fig-year-fowl) provide an overview of how the environmental conditions at Toronto beaches have evolved over the years. Rather than focusing on short-term seasonal changes, these visualizations highlight potential long-term trends and environmental shifts that could have significant implications for beach management and ecosystem health.

## Gain from Seasonal Changes of Beaches

@fig-month-air illustrates the relationship between air temperature and time, focusing on the months from May to October. A clear seasonal pattern emerges, with air temperatures gradually increasing from May, peaking around July, and then declining into September. The seasonal cycle is evident, as air temperatures warm up in late spring, reach their highest points in midsummer, and cool down as fall approaches. This reinforces the influence of seasonal climate variations on beach conditions.

@fig-month-water presents air temperature versus time over the same period. It reveals a similar pattern to air temperature, with a peak in August, which is later than air temperature. This might be due to the absorption of heat in water, causing the temperature to fall gradually. This trend aligns with typical summer weather patterns, where water temperatures rise due to prolonged exposure to sunlight and warmer air, creating optimal conditions for beachgoers.

@fig-month-fowl depicts waterfowl observations versus months from May to September. It shows varying levels of bird activity throughout the summer months, with peaks in July or August. These fluctuations could be influenced by breeding seasons, availability of food, or migratory behavior, all of which are affected by seasonal changes. The variability in bird counts reflects how environmental factors and beach conditions interact with wildlife throughout the warmer months.

In summary, these scatter plots (@fig-month-air, @fig-month-water, and @fig-month-fowl) highlight the clear seasonal trends in both temperature and waterfowl observations at Toronto beaches between May and September. These patterns are important for understanding how environmental conditions shift during the beach season and inform management strategies for maintaining water quality and ecosystem balance during peak usage times.

## Gain from Correlations Between Observed Variables

@fig-air-water depicts the relationship between air temperature and water temperature. It shows a clear positive correlation between the two variables: as air temperatures rise, water temperatures also tend to increase. This relationship is expected, as higher air temperatures lead to more intense solar radiation and warming of the water. However, the variance of water temperature given same air temperature is large. There might be other confounders affecting this relationship. Also, this might be due to the slower rate at which water heats and cools compared to air.

@fig-water-fowl illustrates the relationship between water temperature and waterfowl observations. It does not show a linear relationship between the variables. Rather, it reveals that waterfowl counts are generally higher when water temperatures are moderate, with fewer birds observed at the extreme ends of the temperature range. This suggests that waterfowl may be more active or present in larger numbers when water conditions are neither too cold nor too warm, possibly reflecting optimal conditions for feeding or habitat usage during these periods.

These scatter plots (@fig-air-water and @fig-water-fowl) help clarify how environmental variables at Toronto beaches are interrelated. The strong correlation between air and water temperatures demonstrates how weather patterns influence beach conditions, while the relationship between water temperature and waterfowl observations provides insights into how wildlife activity is shaped by environmental factors. Understanding these connections is crucial for managing beach environments and ensuring both human and ecological health.

# Discussion {#sec-discussion}

## Insights and Implications

The findings (@sec-results) from this analysis of Toronto beach data provide critical insights into the environmental dynamics affecting beach conditions and wildlife interactions. The histograms (@fig-air, @fig-water, and @fig-fowl) illustrate distinct patterns among the measured variables, which can inform both public health monitoring and beach management strategies.

One notable discovery is the temperature differential between air and water, highlighted by the observation that water cools down slower than air. This is evident from the left-skewed distribution of the scatter plot of water temperatures (@fig-month-water), which indicates that even as air temperatures peak in the summer months, water temperatures tend to remain cooler for extended periods. This phenomenon has significant implications for beachgoers, as the cooler water can influence recreational activities and safety protocols. The gradual rise in water temperature, peaking later than air temperature, underscores the thermal inertia of water bodies and emphasizes the importance of considering both air and water temperatures when assessing beach conditions.

The analysis of long-term trends in both air and water temperatures indicates a potential shift in climatic conditions over the years. The scatter plots (@fig-year-air, @fig-year-water, and @fig-year-fowl) reveal that air temperatures exhibit a gradual increase, suggesting that climate change may be impacting local weather patterns. This upward trend in air temperatures may lead to longer and warmer beach seasons, influencing both human behavior and ecological balance. Furthermore, the fluctuating patterns in water temperatures, albeit less explicit, may point toward long-term environmental shifts that require further investigation.

Moreover, the scatter plots (@fig-month-air, @fig-month-water, and @fig-month-fowl) demonstrate seasonal variations in environmental variables, particularly between May and September. The clear patterns observed in air and water temperatures (@fig-air-water) reveal that higher temperatures correspond with peak beach activity, which is vital for planning public health initiatives and environmental monitoring during the warmer months. For example, understanding these seasonal trends allows for the implementation of effective water quality assessments, as increased beach usage during warmer weather may elevate the risk of contamination, especially in the presence of high waterfowl populations.

The relationship between water temperature and waterfowl observations (@fig-water-fowl) also offers valuable insights. The right-skewed distribution of waterfowl counts (@fig-fowl) suggests that while large gatherings of birds are uncommon, they do occur under certain conditions. This variability in bird activity can impact water quality, as large groups of waterfowl can contribute to nutrient loading and potential contamination of water sources. Monitoring these fluctuations is crucial for managing ecosystems, particularly during peak beach seasons when both human activity and wildlife presence are heightened.

These findings point to the need for ongoing research and monitoring efforts. Understanding the interactions between air and water temperatures, along with the behavior of waterfowl, can inform management practices aimed at preserving water quality and ensuring safe recreational spaces for the public. Additionally, the insights gained from this study can guide future investigations into the impacts of climate change on local ecosystems, particularly as shifting weather patterns may alter the historical norms observed in this analysis.

In conclusion, the data analyzed in this paper (@sec-data) highlights the complex interplay between environmental variables at Toronto beaches. Recognizing these patterns not only aids in immediate beach management efforts but also contributes to a broader understanding of how climate dynamics affect local ecosystems. By leveraging these insights, policymakers and environmental managers can better safeguard the health of both human populations and the natural habitats that support them.

## Weaknesses and Next Steps

While this analysis provides valuable insights into the environmental conditions at Toronto beaches, there are several weaknesses that warrant consideration. One major limitation is that the dataset exclusively covers the summer months, specifically from May to September. This narrow time frame restricts our understanding of seasonal variations and trends that occur outside of the summer period, such as the colder months when beach activity is minimal. As a result, the analysis may overlook significant ecological and environmental changes that take place in the winter and spring, including water quality fluctuations, temperature changes, and wildlife patterns.

To address this limitation, future research should aim to incorporate a more comprehensive dataset that includes data from the entire year. This expanded dataset would allow for a more robust analysis of seasonal trends and their implications on beach conditions, wildlife interactions, and public health. Additionally, analyzing data from various seasons could provide insights into how climate change is affecting not only summer conditions but also the overall health and sustainability of the beach ecosystem throughout the year (@beachecosystem).

Another limitation is the presence of missing values (NA values) in the dataset. These NA values can arise from various factors, such as incomplete data collection or reporting errors, and they may lead to biased results if not handled properly. To improve the dataset's robustness, it is essential to implement strategies for dealing with these missing values. Possible approaches include employing statistical techniques such as imputation, where missing values are estimated based on the available data, or conducting sensitivity analyses to understand how the absence of certain data points might influence the overall results (@missingvalue). By addressing NA values more effectively, we can enhance the reliability and validity of the findings.

Furthermore, conducting long-term studies that track changes over multiple years would provide a clearer picture of the trends and shifts in environmental conditions at Toronto beaches. This longitudinal approach could help identify correlations between changing climate patterns and alterations in both air and water temperatures, as well as the behaviors of local wildlife (@climatebeach).

In summary, while this study highlights key patterns and trends in summer beach conditions, expanding the dataset to encompass a full annual cycle and incorporating additional variables will significantly enhance the depth and applicability of future research. These next steps are crucial for developing effective management strategies aimed at preserving the ecological integrity of Toronto beaches and ensuring they remain safe and enjoyable for all visitors.

@@

## Graph and Sketches

Sketches depicting both the desired dataset and the graphs generated in this analysis are available in the GitHub Repository.

@@

The data cleaning process involved renaming variables, tidying the dates, filtering out useless columns from the raw dataset, filtering out observations with NAs, and removing the outliers using z-scores.

@@

```{r}
#| label: tbl-summarise-air
#| tbl-cap: Summarise Statistics of Air Temperatures
#| echo: false
#| warning: false
#| message: false

summarise_air_temp <- data |>
  summarise(Mean = mean(air_temp),
            Median = median(air_temp),
            Variance = var(air_temp),
            Min = min(air_temp),
            Max = max(air_temp),
            IQR = IQR(air_temp)
  )
kable(summarise_air_temp, format = "markdown")
```

```{r}
#| label: tbl-summarise-water
#| tbl-cap: Summarise Statistics of Water Temperatures
#| echo: false
#| warning: false
#| message: false

summarise_water_temp <- data |>
  summarise(Mean = mean(water_temp),
            Median = median(water_temp),
            Variance = var(water_temp),
            Min = min(water_temp),
            Max = max(water_temp),
            IQR = IQR(water_temp)
  )
kable(summarise_water_temp, format = "markdown")
```

```{r}
#| label: tbl-summarise-fowl
#| tbl-cap: Summarise Statistics of Waterfowl Observations
#| echo: false
#| warning: false
#| message: false

summarise_water_fowl <- data |>
  summarise(Mean = mean(water_fowl),
            Median = median(water_fowl),
            Variance = var(water_fowl),
            Min = min(water_fowl),
            Max = max(water_fowl),
            IQR = IQR(water_fowl)
  )
kable(summarise_water_fowl, format = "markdown")
```

>>>>

---
title: "Spotify Data"
author: "Yanzun Jiang, "
date: today
date-format: long
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

To conduct the analysis of songs on Spotify, a dataset downloaded from @spotify was utilized, as described in @sec-data. Based on the initial findings, it was observed that higher air temperatures are correlated with higher water temperatures, while higher water temperatures coincided with an increase in waterfowl activity in some extent(@sec-results). Also, supplementary insights are provided in @sec-appendix.

@@

The dataset used in this analysis is the songs of David Tao sourced from Spotify (@spotify). It records various aspects of the songs by David from 1997 to present.

The variables analyzed in this study are listed below:

`album_release_date`: the date that the album was released

`energy`: the energy of the song

`key`: the main key of the song

`valence`: the degree of happiness of the song

`tempo`: the pace of the song

The dataset was accessed using the `spotifyr` package (@spotifyr). For the analysis, the R programming language was employed (@r), utilizing the `tidyverse` (@tidyverse) package for data cleaning, transformation, visualization, and the `here` package (@here) accessing the data in this analysis. Afterward, the cleaned dataset was processed and tested using additional functions from the `tidyverse` package (@tidyverse).

@@

@@

@@

@@

The data cleaning process involved tidying the dates, filtering out useless columns from the raw dataset, and filtering out observations with NAs.

@@
